The prediction of the corrosion rate in industrial assets relies on
imbalanced databases, in which the observations concentrate at low
values of the target variable while the regimes of higher severity
remain under-represented, and on databases with a reduced number of
useful records. Conventional regression models underestimate the
critical conditions and compromise integrity decisions. This research
proposes a hybrid strategy that articulates segmentation by normative
ranges, interpretable classification, and fuzzy inference, with the
purpose of obtaining a reliable and auditable numerical estimator from
the knowledge extracted from the data and from expert judgment. The
methodology transforms the target variable into the classes defined by
the technical standard: 0--2 [mpy], 2--8 [mpy], and above 8 [mpy]. One
binomial decision tree per normative limit is trained under stratified
nested cross-validation, adjusting the hyperparameters with the
criterion of exceeding a mean per-class recall of 0.6. A record-level
leakage audit precedes the modeling: it revealed that a calculated
thickness-loss column reproduced the reported target exactly for
76.5\% of the records and excluded it together with its ingredients.
On the audited internal-corrosion dataset of 897 records, the trees
reached out-of-fold mean recalls of 0.71 and 0.81 at the 2 [mpy] and
8 [mpy] limits with balanced per-class sensitivity. The rule
extraction produced twelve interpretable propositions, consistent with
the service history and the operating conditions of the asset. Each
numerical partition translates into an interval membership function
anchored at the tree thresholds, and each branch reformulates as a
fuzzy rule whose consequent remains bounded to its normative range.
The fuzzy system converts these discrete rules into a continuous
output, smoothing the transitions between ranges: it reached an
out-of-fold coefficient of determination of 0.50 and a mean absolute
error of 3.6 [mpy], against 0.49 and 4.1 [mpy] for linear regression
on identical folds, with a lower error in every severity zone. The
traceability between each prediction, the active rule, and the
integrity meaning of its conditions supports expert audit of the
complete inference chain in scenarios of imbalanced and reduced data,
frequent in asset-integrity applications.
